---
title: "BSSR_Analysis"
author: "DakaraiMcCoy"
date: "2/25/2020"
output: html_document
---

## Gather all measured variables for each subject into one sheet
```{r load data}
# Objective of this block is to get the common subjects across both waves and all their covariates
# combine all spread sheets onto one page
# check which subjects common to both time points
# rename any variables as necessary
# save a *.csv file that has subjects at both timepoints and denote the differenece with a 'space', '.', '-', or '_' folllowed by a 1 or a 2 extension

```

## initialization
```{r bssr}
# Load a data set that has and one that has not been smoothed.

library(bssr)
# load the data
wave_1 <- load_bss_data(type = "tbm",
                       subjdir = "/Volumes/SocialDevelopmentProject/fMRI/BrainSuite/HandEdited/wave1",
                       csv =   "/Volumes/SocialDevelopmentProject/fMRI/BrainSuite/Stats/Common_subjects.csv",
                       smooth = 0,
                       atlas = "/Applications/BrainSuite19b/svreg/BCI-DNI_brain_atlas/BCI-DNI_brain")

wave_2 <- load_bss_data(type = "tbm",
                        subjdir = "/Volumes/SocialDevelopmentProject/fMRI/BrainSuite/HandEdited/BrainSuite_Wave2_Processed",
                        csv = "/Volumes/SocialDevelopmentProject/fMRI/BrainSuite/Stats/Common_subjects.csv",
                        smooth = 0,
                        atlas = "/Applications/BrainSuite19b/svreg/BCI-DNI_brain_atlas/BCI-DNI_brain")
```
## Whole brain modeling 
The goal here is to explore the main effects of age, sex, culture, IQ and SES.
Functions bss_anova and lm_vec do not allow for modeling without any covariates.

Research Questions:

1. Are there sex differences between adolescent brains at each wave? If so, how are the changes related to the imaging timepoints?   
  OBSERVATIONS: Puberty is a time during which a cascade of horomones are used to signal the body it is time to develop 
                reproductive organs. These signals lead to changes in the brain that affect its organization and
                activity. Studies have shown the male brain is about 10% larger in volume than female brains when
                controlling for body size and weight( 0 ). It has been shown that Recent studies have also begun to
                investigate d
  HYPOTHESIS:
  TESTABLE PREDICTION:
  RESULTS:
  INTERPRETATION:
  CONCLUSION:

2. Is there a main effect of culture in the adolescent brains at each wave? Does the effect affect adolescent brain development across timepoints? 
  OBSERVATIONS:
  HYPOTHESIS:
  TESTABLE PREDICTION:
  RESULTS:
  INTERPRETATION:
  CONCLUSION:

3. Is there a main effect for the change in age (wave2 - wave1) of the adolescent brain development? 
  OBSERVATIONS:
  HYPOTHESIS:
  TESTABLE PREDICTION:
  RESULTS:
  INTERPRETATION:
  CONCLUSION:
  
4. Is there a main effect for the type and length of violence exposure? In what brain regions? And does this effect hold when considering other covariates?
  OBSERVATIONS:
  HYPOTHESIS:
  TESTABLE PREDICTION:
  RESULTS:
  INTERPRETATION:
  CONCLUSION:

---> Follow up with Shantanu about this 2/28/20


```{r tbm_wave1}
# Here I have noticed I can run paired ttest to test for a difference across timepoints for each subject! 

# Test for the main effect of sex for subjects in wave 1. 
sex_ttest_1 <- bss_ttest(group_var = "sex", bss_data = wave_1, paired = FALSE)

#Test for the main effect of culture for subjects in wave 1
culture_main <- bss_ttest(group_var = "culture", bss_data = wave_1, paired = FALSE)

# test for the main effect of IQ for subjects in wave 1
age_ttest <- bss_ttest(group_var = "age at wave 1", bss_data = wave_1, paired = FALSE)

### NEED TO GET DATA FROM FEI ON VIOLENCE EXPOSURE ---> 2/28/20
# # Test for the main effect of violence exposure
violence_ttest <- bss_ttest(groupvar = "violence", bss_data = wave_1, paired = FALSE)


# If I find a significant whole brain effect, I want to know if the main effect will remain when including different covariates

# Test for the main effect of sex when controlled by the age at wave 1. 
sex_model <- bss_anova(main_effect = "sex", covariates = "age at wave 1", bss_data = wave_1, mult_comp = "fdr" )

# Test fot the main effect of culture when controlled by the age at wave 1.
culture_model <- bss_anova(main_effect = "culture", covariates = "NULL", bss_data = wave_1, mult_comp = "fdr")

# Test the difference between lm_vec and the the bss_anova without any covariates. Are the results the same? 
sex_model_regression <- lm_vec(main_effect = "sex", covariates = "age at wave 1", bss_data = wave_1)




```

```{r save data/model}
save_bss_out(bss_data = wave_1, sex_ttest_1, outdir="~/Desktop/tbm_anova/sex")
save_bss_out(wave_1, , outdir = "~/Desktop/tbm_anova/lm_vec")
save_bss_out(wave_1, age_ttest, outdir = "~/Desktop/tbm_anova_sex/ttest_1")
```