---
title: "Structural Descriptive Statistics"
author: "DakaraiMcCoy"
date: "8/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# Need R 4.0 to get psych toolbox to work. 
# library(psych)
```

## Get Descriptive Statistics - Method validation
I need to check which subjects are outliers and which subjects are outliers in each region. It appears as though there may be a functional-structure relationship in these areas of the brain. Questions from lab meeting are: Are these brain regions actually measures we can use? Are they representative of the population and can we say these range of values is acceptable for developing adolescents? 

## Read Participant information excel sheet and format to join with other data types. 
```{r participant info}

# Load participant information
ses_df <- readxl::read_xlsx("Participant_info.xlsx")

# Get subjectsbb that are in both timepoints
subject_to_load <- ses_df %>% 
  mutate(wave1_data = file.exists(paste0("~/Shared/BrainSuite/Stats/wave1/sub", `Subject #`, "-anat.roiwise.stats.txt")), 
         wave2_data = file.exists(paste0("~/Shared/BrainSuite/Stats/wave2/", `Subject #`, "_T1w_1m.roiwise.stats.txt"))) %>% 
  filter(wave1_data, wave2_data) %>% 
  pull(`Subject #`) 

# Get information for subjects that are in both timepoints
participant_info <- ses_df %>% 
  mutate(wave1_data = file.exists(paste0("~/Shared/BrainSuite/Stats/wave1/sub", `Subject #`, "-anat.roiwise.stats.txt")), 
         wave2_data = file.exists(paste0("~/Shared/BrainSuite/Stats/wave2/", `Subject #`, "_T1w_1m.roiwise.stats.txt"))) %>% 
  filter(wave1_data, wave2_data) %>%
  select(subject = `Subject #`, culture = starts_with("culture"), 
         gender = starts_with("gender"), starts_with("age")) %>%
  rename(age_1 = "age at wave 1", age_2 = "age at wave 2")

  
```


```{r merged data set}
# This function loads subjects' BrainSuite statisics file using paste0 to create the fullpath for each subject. Function uses map_df to apply the read.delim function to each file, creating a long format of data. 
load_subjects <- function (subjectID, path, extension) {
  map_df(subjectID, 
         ~ read.delim(paste0(path, .x, extension)) %>% 
           mutate(subject = .x))
}

# Function to get inusla regions and other BrainSuite measurements
# TO DO: Get a general function to map ROI ~ (hemisphere | region)

get_Insula_common_subs <- function (x, time) {
  x %>%
  # Get the total brain volume for each suject
  mutate(total_Vol = GM_Volume.mm.3. + WM_Volume.mm.3. + CSF_Volume.mm.3.) %>%
  # Get hemisphere for ROI
  mutate(hemisphere = ifelse( ROI_ID %in% c("500", "510", "520", "530"), 
                            "right", "left")) %>%
  # Save only variables I need
  select(subject, ROI_ID, Mean_Thickness.mm., GM_Volume.mm.3., 
         WM_Volume.mm.3., CSF_Volume.mm.3., total_Vol, hemisphere ) %>% 
  # subset dataframe by ROI_ID
  filter(ROI_ID >= 500 & ROI_ID < 600) %>%
  # add region to the data frame for ROI
  mutate(region = factor(substr(ROI_ID, 1, 2),
                        labels = c("In", "dAI", "vAI", "postI"))) %>%
  # scale brain measures by total volume in each region
  mutate_at(c("GM_Volume.mm.3.", "WM_Volume.mm.3.", "CSF_Volume.mm.3."), 
            ~./total_Vol) %>%
  # Drop ROI and rename measurement variables. 
  select(subject, hemisphere, region, 
         MT = Mean_Thickness.mm., GM = GM_Volume.mm.3., 
         WM =  WM_Volume.mm.3., CSF = CSF_Volume.mm.3., TV = total_Vol) %>%
  gather(key = "var", value = "value", 4:8) %>% 
  unite("var_wide", hemisphere, region, var) %>% 
  spread(var_wide, value) %>%
    # Get normalized total inusla volume in
  mutate(left_In = left_dAI_GM + left_dAI_WM + 
                   left_vAI_GM + left_vAI_WM + 
                   left_postI_GM + left_postI_WM + 
                   left_In_GM + left_In_WM) %>%
  mutate(right_In = right_dAI_GM + right_dAI_WM +  
                   right_vAI_GM + right_vAI_WM  +
                   right_postI_GM + right_postI_WM + 
                   right_In_GM + right_In_WM) %>%
  mutate(time = time)
}

# Are there outliers inthe total brain volume? Does it make sense to have outliers? Are the outliers interpretable? If the two brains are outliers? Are they actually larger when comparing side by side when the aspect ration is 1:1, use BrainSuite viewing tool? Is the difference in size due to an artifact? 

# Does it make sense to normalize mean thickness with total brain volume.
# Literature on mean thickness... Is it typically normalized by total brain volume?


# Load the subjects
wave_1 <- load_subjects(subject_to_load,
                        "~/Shared/BrainSuite/Stats/wave1/sub",
                        "-anat.roiwise.stats.txt")

wave_2 <- load_subjects(subject_to_load,
                        "~/Shared/BrainSuite/Stats/wave2/",
                        "_T1w_1m.roiwise.stats.txt")

# Get Insula from common subjects
wave1_wide_normalize <- get_Insula_common_subs(wave_1, 1)
wave2_wide_normalize <- get_Insula_common_subs(wave_2, 2)

# Create wide data format for both timepoints
# Add parcipant information to the merged data set
merged <- merge(wave1_wide_normalize, wave2_wide_normalize, all = TRUE) %>% 
  gather(key = "var", value = "value" , 2:43) %>% 
  unite("var_wide",var,time) %>% 
  spread(var_wide,value) %>%
  merge(participant_info)


#psych::describe(merged)
#psych::pairs.panels(merged %>% select(left_dAI_GM_1, age_1,age_2,culture, gender))

```

```{r boxplot_wave1}
# Visual inspection of subjects. 
# Show boxplots
library(ggplot2)
#p <- ggplot() + 
  


csf_1 = boxplot(ylab = "normalized Volume a.u.", 
             wave1_wide_normalize[seq(from=2, to=dim(wave1_wide_normalize)[2], by=4)])
gM_1 = boxplot(ylab = "normalized Volume a.u.", 
             wave1_wide_normalize[seq(from=3, to=dim(wave1_wide_normalize)[2], by=4)])
mT_1 = boxplot(ylab = "units in mm^2", 
             wave1_wide_normalize[seq(from=4, to=dim(wave1_wide_normalize)[2], by=4)])
wM_1 = boxplot(ylab = "normalized Volume a.u.", 
             wave1_wide_normalize[seq(from=5, to=dim(wave1_wide_normalize)[2], by=4)])

# Now add wave 2 to the data set.
  

```

# Get outliers for same regions as before with normalized brain volumes
```{r normalized outliers}

# Doesnt quite work yet. But it should work
# get_outlier_subjects <- function(data,outliers) {
#   out <- c()
#   for (i in 1:ncol(data)){
#     out <- as.list(out,data.frame(outliers_i = match(signif(outliers$out[outliers$group == i], 7),
#                                                        signif(data[,i], 7))))
#     
#     out <- rbind(out,
#                  subs <- data.frame(outliers_i = match(signif(outliers$out[outliers$group == i], 7),
#                                                        signif(data[,i], 7))))
#   }
#   }
# Turn this into a function later on
# 

# Gm <- get_outlier_subjects(gm,gm.boxplot)
# Mt <- get_outlier_subjects(mt,mt.boxplot)
# Wm <- get_outlier_subjects(wm,wm.boxplot)

# outliers for MT
# GM <- c()
# for (i in 1:ncol(gm)){
# GM <- rbind(GM,
#               sub1 <- data.frame(outliers_ind = match(signif(gm.boxplot$out[gm.boxplot$group == i], 7),
#                                                        signif(gm[,i], 7))))
# }
# GM$group = gm.boxplot$group

# matcher <- function (boxObj,data){
#   hits = tibble::tibble(.rows = 2)
#   for(i in seq(from=1, to=length(boxObj$group), by=1)){
#          hits[[1,i]] = match(signif(boxObj$out[boxObj$group == i], 7),
#                       signif(data[,i], 7))[which.min(boxObj$out[boxObj$group == i])]
#          hits[[2,i]] = match(signif(boxObj$out[boxObj$group == i], 7),
#                       signif(data[,i], 7))[which.max(boxObj$out[boxObj$group == i])]
#          }
#   return(hits)
#   }
# try map_df function
# map_df(1:8, function (i) {match(signif(gM_1$out[gM_1$group == i], 7),
#                      signif(data[,i], 7))[which.max(gM_1$out[gM_1$group == i])]})

# Find the max and min of each measure in all Insula regions.

# Group 7 ROI - Right post Insula 
rPI_min <-match(signif(gM_1$out[gM_1$group == 7], 7),
                signif(wave1_wide_normalize[seq(from=3,                                    to=dim(wave1_wide_normalize)[2],by=4)][,7],7))[which.min(gM_1$out[gM_1$group ==7])]
rPI_max <- match(signif(gM_1$out[gM_1$group == 7], 7),
      signif(wave1_wide_normalize[seq(from=3,                                    to=dim(wave1_wide_normalize)[2],by=4)][,7], 7))[which.max(gM_1$out[gM_1$group ==7])]

# GM Group 3 outliers - left post Insula
lPI_min <- match(signif(gM_1$out[gM_1$group == 3], 7),
      signif(wave1_wide_normalize[seq(from=3,                                    to=dim(wave1_wide_normalize)[2],by=4)][,3], 7))[which.min(gM_1$out[gM_1$group ==3])]
lPI_max <- match(signif(gM_1$out[gM_1$group == 3], 7),
      signif(wave1_wide_normalize[seq(from=3,                                    to=dim(wave1_wide_normalize)[2],by=4)][,3], 7))[which.max(gM_1$out[gM_1$group ==3])]

# GM Group 1 outliers - left dorsal anterior Insula
ldAI_min <- match(signif(gM_1$out[gM_1$group == 1], 7),
      signif(wave1_wide_normalize[seq(from=3,                                    to=dim(wave1_wide_normalize)[2],by=4)][,1], 7))[which.min(gM_1$out[gM_1$group ==1])]
ldAI_max <- match(signif(gM_1$out[gM_1$group == 1], 7),
      signif(wave1_wide_normalize[seq(from=3,                                    to=dim(wave1_wide_normalize)[2],by=4)][,1], 7))[which.max(gM_1$out[gM_1$group ==1])]

# GM Group 5 outliers - right dorsal anterior Insula
rdAI_min <- match(signif(gM_1$out[gM_1$group == 5], 7),
      signif(wave1_wide_normalize[seq(from=3,                                    to=dim(wave1_wide_normalize)[2],by=4)][,5], 7))[which.min(gM_1$out[gM_1$group ==5])]
rdAI_max <- match(signif(gM_1$out[gM_1$group == 5], 7),
      signif(wave1_wide_normalize[seq(from=3,                                    to=dim(wave1_wide_normalize)[2],by=4)][,5], 7))[which.max(gM_1$out[gM_1$group ==5])]

# GM Group 2 outliers - left Insula
lI_min_lI <- match(signif(gM_1$out[gM_1$group == 2], 7),
      signif(wave1_wide_normalize[seq(from=3,                                    to=dim(wave1_wide_normalize)[2],by=4)][,2], 7))[which.min(gM_1$out[gM_1$group == 2])]
lI_max <- match(signif(gM_1$out[gM_1$group == 2], 7),
      signif(wave1_wide_normalize[seq(from=3,                                    to=dim(wave1_wide_normalize)[2],by=4)][,2], 7))[which.max(gM_1$out[gM_1$group == 2])]


```

```{r Model_1}
## Null Hypothesis is that there is no group difference in mean cortical thickness. alternative Hypothesis is that there is a group difference in mean cortical thickness. 

```

## Plot statistics using sex, and age to see trends
```{r}
# Plot the ROI's by age and sex. 
  

```

## Diagnostic tools to check if all assumptions are met for each type of  

