---
title: "Structural Descriptive Statistics"
author: "DakaraiMcCoy"
date: "8/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import data
```{r import data}
# Use R to get a list of files 
temp <- list.files("/Volumes/SocialDevelopmentProject/fMRI/BrainSuite/Automated/wave2", pattern = "*.txt", full.names = TRUE)
# Create list of data frames
myfiles = lapply(temp,read.delim)

```

## Clean data
```{r}
library(plyr)
       
# In each data frame, create a subset data frame for the ROI and the measurments of interest. 
myfiles.ROI <-lapply(myfiles, function(x) {subset.data.frame(x,ROI_ID >= 500 & ROI_ID < 600,c(ROI_ID,Mean_Thickness.mm.,GM_Volume.mm.3.,WM_Volume.mm.3.))})

# Melt the data to create a long data frame suitable for casting 
library(reshape2)

myfiles.long <- melt(myfiles.ROI, id.vars = c("Mean_Thickness.mm.", "GM_Volume.mm.3.", "WM_Volume.mm.3."), measure.vars = "ROI_ID", value.name = "ROI")

# Make sure the ROI's are categrical.
myfiles.long$ROI <- myfiles.long$ROI <- as.factor(myfiles.long$ROI)

# Cast ROI's as categorical variables
#str(myfiles.long)

```

## Read each dataframe and get averages
```{r}



```
## Plot statistics using sex, and age to see trends


## Get Descriptive Statistics
## using non-parametric pertubations, is there a group difference in wave 1. 
## Null Hypothesis is that there is no group difference in mean cortical thickness. alternative Hypothesis is that there is a group difference in mean cortical thickness. 

## Null Hypothesis is that there is no difference of the mean cortical thickness between the two groups of students. Alternative hypothesis is that there is a difference of the mean cortical thickness between the two groups of students. 
```{r}
library(ggplot2)

# Show box plot of mean cortical thickness in each ROI
p <- ggplot(myfiles.long, mapping = aes(x = "Mean Thickness in mm", y = Mean_Thickness.mm.)) +
             geom_boxplot() +
             geom_jitter(width = 0.2) +   # add a plot layer with points
             facet_wrap(~ROI, ncol = 2)
show(p)

# Show box plot of WM Volume in each ROI
p <- ggplot(myfiles.long, mapping = aes(x = "White Matter Volume in mm^3", y = myfiles.long$WM_Volume.mm.3.)) +
             geom_boxplot() +
             geom_jitter(width = 0.2) +   # add a plot layer with points
             facet_wrap(~ROI, ncol = 2)
show(p)

# Show box plot of GM Volume in each ROI
p <- ggplot(myfiles.long, mapping = aes(x = "Gray matter Volume in mm^3", y = myfiles.long$GM_Volume.mm.3.)) +
             geom_boxplot() + 
             #geom_jitter(width = 0.2) +   # add a plot layer with points
             facet_wrap(~ROI, ncol = 2)
show(p)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
